---
title: "Files and Folders"
output: html_notebook
---

# About this R Notebook

R Notebooks are a 'flavor' of R markdown that combine plain text with R commands in code chunks. You can (and should!) edit them, and save it every few minutes like anything else!

Minimize the console. You won't need it because the output will appear 

To view a copy of your notebook as HTML, just hit Save, then Preview.

Keyboard shortcuts:

 - run the current line of R: *ctrl + enter*
 - run everything in the current code chunk: *ctrl + shift + enter*
 - insert a new code chunk: *ctrl + alt + i*

# Seeing what files exist

View the working directory

```{r}
getwd()
```
List the files in the current working directory. Note '.' is shorthand for the current working directory.

```{r}
list.files(".")
```

**Challenge**. List all the files in your 'Home' directory. [Answer](http://bit.ly/38UPrbQ) 

```{r}

```

List the Shapefiles in the 'data' directory.

```{r}
list.files("./data", pattern = "*.shp")
```
Write an expression that will check if a file named "sf_libraries.csv" exists in the data directory.

```{r}
file.exists("./data/sf_libraries.csv")
```

Import the *sf_libraries.csv* file. 

```{r}
my_data <- read.csv("./data/sf_libraries.csv")

# View(my_data)
```

**Challenge**: What type of R object does `read.csv()` return?

```{r}

```

Create 1000 random values using the random number function of your choice (e.g., rnorm).

```{r}
rnd_vals <- rnorm(1000) * 20
```

**Challenge**: Create a histogram of the values.

```{r}

```

Save your random numbers to your 'Home' directory.

```{r}
save(rnd_vals, file="~/my_random_numbers.RData")
```

# End

Congratulations on finishing the Notebook! Save it (again).

Click on the 'Preview' button in the RStudio toolbar to save a copy of your work as HTML.


