---
title: "Wrangling 2D Data with dplyr"
output: 
  html_notebook:
    toc: yes
    toc_float: yes
---

`dplyr` is a package for manipulating data frames, including subsetting rows and columns, sorting rows, adding new columns, grouping rows, joins, etc.

In this notebook we'll practice using the `dplyr` package.

## Set preferences with `conflicted`

Several `dplyr` functions have very generic names. To avoid potential conflicts with functions from  other packages, we'll start by stating our preferences using the conflicted package:

```{r chunk01}
library(dplyr)

## Load the conflicted package
library(conflicted)

# Set conflict preferences
conflict_prefer("filter", "dplyr", quiet = TRUE)
conflict_prefer("count", "dplyr", quiet = TRUE)
conflict_prefer("select", "dplyr", quiet = TRUE)
conflict_prefer("arrange", "dplyr", quiet = TRUE)
```

Remember the iris data frame? How many rows does iris have? 

```{r chunk02}
nrow(iris)
```

## Filter rows with `filter()`

Write an expression that returns just those rows where Sepal.Length > 7:

```{r chunk03}
# Select rows where Sepal.Length > 7
iris %>% filter(Sepal.Length > 7)
```

Add another criteria specifying the species has to be `virginica`:

HINT: remember to use '==' to test for equality

```{r chunk04}
iris %>% filter(Sepal.Length > 7, Species=="virginica")
```

**CHALLENGE**: Write an expression that will return rows where the Petal.Length is between 4.1 and 4.4. [Answer](https://bit.ly/3tNPVIH)

```{r chunk05}

```

## Filter rows with `slice()`

Sometimes you want to fliter rows by row number, rather than an attribute column. You can do that with `slice()`.

Show the first 10 rows from iris:

```{r chunk06}
iris %>% slice(1:10)
```

## Filter columns with `select()`

To get a subset of columns, or if you want the columns in a different order, include the `select()` function in your expression.

For example if you only want the `Sepal.Length` and `Sepal.Width` columns, you could run:

```{r chunk07}
iris %>% 
  filter(Sepal.Length > 7) %>% 
  select(Sepal.Length, Sepal.Width, Species)
```

## Sort rows with `arrange()`

Add `arrange()` to your expression to sort rows. For example to sort the last expression by `Sepal.Width` we simply add one more piece to the expression:

```{r chunk08a}
iris %>% 
  filter(Sepal.Length > 7) %>% 
  select(Sepal.Length, Sepal.Width, Species) %>% 
  arrange(Sepal.Width)
```

To sort rows in descending order, wrap the field name in `desc()`:

```{r chunk08b}
iris %>% 
  filter(Sepal.Length > 7) %>% 
  select(Sepal.Length, Sepal.Width, Species) %>% 
  arrange(desc(Sepal.Width))
```

**CHALLENGE**: Write an expression that returns the top-5 Petal Lengths (all species)

Hint: there are a couple of ways to do this. Look at the dplyr [Cheatsheet](https://github.com/rstudio/cheatsheets/raw/master/data-transformation.pdf) for inspiration. [Answer](https://bit.ly/3d9rVsS)

```{r chunk08c}

```

## Create a new column with `mutate()`

Create a new column which contains the product of the Sepal Length and Petal Length. 

```{r chunk09}
iris %>% mutate(sepal_petal_product = Sepal.Length * Petal.Length)
```

**CHALLENGE**: Write an expression from the `quakes` data frame that:

- returns the lat, long, and mag columns  
- only return quakes that are i) north of 20 degrees South, and ii) larger than magnitude 5.0. 
- sorts the rows by magnitude

Hint: Fiji is south of the equator. [Answer](https://bit.ly/3fdflLR)

```{r chunk10}

```

## Group and Summarize Rows

You can 'group' rows using `group_by()`. By itself, grouping rows doesn't do much. Usually you follow `group_by()` with summarise() which creates some kind of summary for each group.

Example: for each Species in iris, compute the mean petal length and width

```{r chunk11}
iris %>% group_by(Species) %>% 
  summarise(avg_petal_length = mean(Petal.Length), avg_petal_width = mean(Petal.Width))
```

**CHALLENGE**: For each species in iris, compute the minimum and maximum sepal length.

```{r chunk12}

```

## Join Tables Based on a Common Column

Practice by creating a data frame with additional info about the three iris species:

```{r chunk13}
iris_colors <- data.frame(Species=c("setosa", "versicolor", "virginica"),
                          num_genes = c(42000, 41000, 43000),
                          max_water = c(1.2, 1.4, 1.7))
iris_colors
```

Now join this table to `iris` using `left_join()`. The `by` argument is the name of the column the data frames have in common.

```{r chunk14}
iris %>% left_join(iris_colors, by = "Species")
```

NOTE: if the common column was named differently in the two tables, you can pass a named vector as the by argument. For example if the capitalization was different you could let `by = c("Species" = "species")`.

## End

Congratulations, you have completed the Notebook! 

To view your Notebook at HTML, save it (again), then click the 'Preview' button in the RStudio toolbar. If


