---
title: "Geoprocessing II"
output: 
  html_notebook: 
    css: nb.css
---

```{r setup}
library(sf)
library(dplyr)
library(units)
library(tibble)
library(tmap)
```

# Dissolve Boundaries Based on Attribute Fields

First, (re)import the watersheds:

```{r}
## Import watersheds
gpkg_watershd_fn <- "./data/yose_watersheds.gpkg"
file.exists(gpkg_watershd_fn)
yose_watersheds <- st_read(gpkg_watershd_fn, layer="calw221")
```
Next, group the rows of the **attribute table** (without geometry) on the 'HU' column (larger watershed ID), and compute the average watershed size.

```{r}
watersheds_df <- yose_watersheds %>% st_drop_geometry()
hu_df <- watersheds_df %>% 
  group_by(HU) %>% 
  summarise(HUNAME = first(HUNAME), 
            NUM_WATERSHEDS = n(),
            AVG_ACRES = mean(ACRES))
head(hu_df)
```

Do the same but this time with the **sf** object (i.e., keeping the geometry). The result is another **sf** object.

```{r}
hu_sf <- yose_watersheds %>% 
  group_by(HU) %>% 
  summarise(HUNAME = first(HUNAME), 
            NUM_WATERSHEDS = n(),
            AVG_ACRES = mean(ACRES))
hu_sf
```
Plot the summarized watersheds:

```{r}
## Plot
tm_shape(hu_sf) + 
  tm_polygons(col="HUNAME", title="HU Watershed") + 
tm_shape(yose_watersheds) +
  tm_borders(col="gray60", lwd=0.8) +
tm_layout(legend.outside = T)
```
# Example: Buffer, Union & Clip Yosemite's Roads

In this example, we’ll create a 500m buffer around the roads within the park (representing a disturbance zone for sensitive species, or drive transect detection area), with the end goal of calculating the percent area of the entire park that lies within 500m of a road. 

Import the roads and the park boundary:

```{r}
## Import the roads
gdb_fn <- "./data/yose_roads.gdb"
yose_roads <- st_read(gdb_fn, "Yosemite_Roads")

## Define a convenience variable for UTM Zone 11
epsg_utm11n_nad83 <- 26911

## Import the YNP border
yose_bnd_utm <- st_read(dsn="./data", layer="yose_boundary") %>% 
  st_transform(epsg_utm11n_nad83)

```

Create the buffer:

```{r}
## Create the buffer
yose_roads_buff <- yose_roads %>% st_buffer(dist=500)

## Plot the results
tm_shape(yose_roads_buff) + tm_polygons(col="red") +
tm_shape(yose_bnd_utm) + tm_borders(lwd=2)
```
Clip to park boundary & union:

```{r}
yose_roads_inpark_buff <- yose_roads_buff %>%
  st_intersection(yose_bnd_utm) %>% 
  st_union()

## Plot the results
tm_shape(yose_bnd_utm) + 
  tm_borders(lwd=2) +
tm_shape(yose_roads_inpark_buff) +
  tm_polygons(col="red")

```
What is the total area of the buffer?

```{r}
road_area <- yose_roads_inpark_buff %>% 
  st_area() %>% 
  sum()
road_area
```
Express the area in ha:

```{r}
road_area %>% set_units(ha)
```
# YOUR TURN!

How big is the buffered roads as a percentage of the total park area?

```{r}
## Road buffer areas as a percentage of the park

```

Answer: 7.9%. See <https://ucanr-igis.github.io/rspatial_scgis20/geoprocessing2.html>

# Example: Voronoi tesselation Yosemite's Campgrounds

First, import the campgrounds (projecting them to UTM Zone 11):

```{r}
## Define a convenience variable for UTM Zone 11
epsg_utm11n_nad83 <- 26911

## Import and project the campgrounds
yose_campgrnds_utm <- st_read(dsn="./data", layer="yose_poi") %>% 
  dplyr::filter(POITYPE == 'Campground') %>% 
  dplyr::select(POINAME) %>% 
  st_transform(epsg_utm11n_nad83)
```

Next, compute the Voronoi polygons:

```{r}
## Compute Voronoi polygons
v <- st_voronoi(st_union(yose_campgrnds_utm))

## Construct polygon sf and simplify (i.e., break multipart into single features)
v_sf <-  st_sf(v) %>% st_cast()

## Plot
plot(v_sf %>% st_geometry() , col = NA, border="gray70")
plot(yose_bnd_utm %>% st_geometry(), add=TRUE, col=NA, border="darkblue")
plot(yose_campgrnds_utm %>% st_geometry(), col="darkgreen", cex=1, pch=17, add=T)
```

# End

Click on the 'Preview' button in the RStudio toolbar to save a copy of your work as HTML.


