---
title: "Create a Leaflet Map from Google Sheets Data"
output: 
  html_notebook:
    toc: yes
    toc_float: yes
---

In this Notebook, we'll make a leaflet map using data from Google Sheets.

## Import San Francisco Public Library Locations from Google Sheets

Start by importing the *sf_libraries* sheet from the following Google Spreadsheet (which can be viewed by anyone with the link):

<https://docs.google.com/spreadsheets/d/1Zm3BkQaW4vuk-X4L_6BOhFEyCr3HO_kSMktd-rHyXG8/>.

\

To import, we're going to need the Google Sheet ID (which you can read from the URL):

```{r chunk01}
gsheet_id <- "1Zm3BkQaW4vuk-X4L_6BOhFEyCr3HO_kSMktd-rHyXG8"
```

\

Load the `googlesheets4` package:

```{r chunk02}
library(googlesheets4)
```

\

Our Google Sheet is open to anyone with the link, so we don't need to be 'signed-in'. To avoid a password prompt, let's 'sign-out' with `gs4_deauth()`:

```{r chunk03}
gs4_deauth()
```


Import the *sf_libraries* sheet using `range_read()`:

```{r chunk04}
## Download Sheet1
sf_libraries_tbl <- range_read(ss = gsheet_id, 
                               sheet = "sf_libraries", 
                               range = NULL)
```

\

View what we got:

```{r chunk05}
tibble::glimpse(sf_libraries_tbl)
```

We can see there's a `lon` and `lat` column in there, so we don't have to geocode the addresses. 

Let's create a version of this tibble containing only the columns we actually need in our interactive map:

```{r chunk06}
library(dplyr, quietly = TRUE)

sf_lib4map_tbl <- sf_libraries_tbl %>% 
  dplyr::select(branch, address, city, state, zip, phone, lon, lat)
sf_lib4map_tbl
```

\

## Create a sf data frame

The next step is turn the data frame into a sf data frame.

The coordinates are already in geographic coordinates (i.e., longitude & latitude), so we don't need to (un)project the points. However we do need to tell sf what the CRS is (4236 is the EPSG number for geographic coordinates with WGS 84 datum).

```{r chunk07}
library(sf, quietly = TRUE)
sf_lib4map_sf <- st_as_sf(sf_lib4map_tbl, coords = c("lon", "lat"), crs = 4326)
sf_lib4map_sf
```

\

## Create an Interactive Map with `tmap`

Load `tmap` and set the mode to 'view' (interactive):

```{r chunk08a}
library(tmap, quietly = TRUE)
tmap_mode("view")
```

\

The most basic tmap involves a `tm_shape()` (to define the data for a layer) plus a function to draw features (like `tm_symbols()`): 

```{r chunk08b}
tm_shape(sf_lib4map_sf) +
  tm_symbols()
```

Let's change the basemap to *OpenStreetMap.* (Note: you can view all of the available basemaps by typing `providers$` in the code chunk or console.)

```{r chunk09}
tm_shape(sf_lib4map_sf) +
  tm_symbols(col = "blue", size = 0.8) +
tm_basemap(providers$OpenStreetMap)
```

\

Lastly we'll modify the pop-up windows with the `popup.vars` argument:

```{r chunk10}
tm_shape(sf_lib4map_sf) +
  tm_symbols(col = "blue", size = 0.8, popup.vars = c("address", "city", "state", "zip", "phone")) +
tm_basemap(providers$OpenStreetMap)
```

\

## Create an Interactive Map with `leaflet`

The `leaflet` package can also be used to make interactive maps. `leaflet` provides a few more customization options, including the ability to customize the pop-up windows with our own HTML text. 

First we create a character vector of HTML code we want to appear in the popup windows:

```{r chunk11}
## Create a character vector of HTML code for the pop-up windows
my_popup_html <- sf_lib4map_sf %>%
  st_drop_geometry() %>% 
  transmute(popup_html = paste0("<b>", branch, "</b><br/>", 
                                address, "<br/>", 
                                city, ", ", state, " ", zip, "<br/>", 
                                "tel: ", phone)) %>% 
  pull(popup_html)

head(my_popup_html)
```

\

Now we're ready to create the map in leaflet:

```{r chunk12}
library(leaflet)

## Create the leaflet map
library_leaf <- leaflet(sf_lib4map_sf) %>% 
  addTiles() %>% 
  addCircleMarkers(popup = my_popup_html, radius = 5, stroke = FALSE, fillOpacity = 0.7)

## Display the leaflet map
library_leaf
```

Click on one of the circles to view the custom HTML code we created for the popup windows

\

## Challenge

**CHALLENGE**: Create a similar interactive map of the CA Breweries (see the *ca_brew* sheet in the same Google Spreadsheet). [Answer](https://bit.ly/3fd3VYC)

```{r chunk13}

```

## End

Congratulations, you've completed another Notebook! 

To view your Notebook at HTML, save it (again), then click the 'Preview' button in the RStudio toolbar.

