---
title: "Import and Map iNaturalist Observations"
output: 
  html_notebook:
    toc: yes
    toc_float: yes
---

In this Notebook, we'll learn how to import observations from [iNaturalist](https://www.inaturalist.org/) using the [rinat](https://docs.ropensci.org/rinat/) package.

## Load Packages

The first step is to load the packages we'll need. These are all on CRAN if you need to install any.

```{r chunk01, message = FALSE}
library(rinat)
library(sf)
library(dplyr)
library(tmap)
library(leaflet)

## Load the conflicted package and set preferences
library(conflicted)
conflict_prefer("filter", "dplyr", quiet = TRUE)
conflict_prefer("count", "dplyr", quiet = TRUE)
conflict_prefer("select", "dplyr", quiet = TRUE)
conflict_prefer("arrange", "dplyr", quiet = TRUE)
```

## Import the Park Boundary

Next, we import the boundary of Yosemite and grab its bounding box. We'll use this later when we call iNaturalist to specify that we're we're only interested in observations in this area.

```{r chunk02}
## Import the YNP Boundary
yose_bnd_ll <- sf::st_read(dsn="./data", layer="yose_boundary")

## Get the bounding box
yose_bnd_bb <- st_bbox(yose_bnd_ll)
as.numeric(yose_bnd_bb)
```

## Retrieve Recent iNat Observations

Next, we can retrieve observations within the bounding box of the park. By default, the `get_inat_obs()` will return the most recent 100 observations:

```{r chunk03}
## Retrieve the first 100 iNaturalist observations within the bounding box. 
## Note we have to rearrange the coordinates of the bounding box a little bit to 
## give get_inat_obs() what it expects

yose_inat_df <- get_inat_obs(bounds = yose_bnd_bb[c(2,1,4,3)])
```

Inspect the results:

```{r chunk04}
dim(yose_inat_df)
head(yose_inat_df)
```

In order to plot these observations, let's first convert the data frame into a sf object:

```{r chunk05}
yose_inat_sf <-  yose_inat_df %>% 
  select(longitude, latitude, datetime, common_name, scientific_name, user_login) %>% 
  st_as_sf(coords=c("longitude", "latitude"),  crs=4326)
yose_inat_sf
```

## Plot with tmap

```{r chunk06}
tmap_mode("view")

tm_shape(yose_bnd_ll) + 
  tm_borders(col = "red", lwd = 2) +
tm_shape(yose_inat_sf) + 
  tm_symbols()
```

## Add a Taxon to the Query 

Next we'll a taxon to our query. Yosemite has some endemic toads. We can tell iNaturalist we only only want certain species using the optional taxon_name argument. Here we'll set it to  [*Bufonidae*](https://amphibiaweb.org/lists/Bufonidae.shtml) (the toad family). 

```{r chunk07}
yose_toads_df <- get_inat_obs(bounds = yose_bnd_bb[c(2,1,4,3)],
                              taxon_name = "Bufonidae")
dim(yose_toads_df)
glimpse(yose_toads_df, width = 110)
```

## Convert the toads df to sf

Before we can plot it, we'll convert the toads layer to a sf object. At the same time, we'll reduce the number of columns to just those we want to include on the map.

```{r chunk08}
yose_toads_sf <-  yose_toads_df %>% 
  select(longitude, latitude, datetime, common_name, scientific_name, image_url, user_login) %>% 
  st_as_sf(coords=c("longitude", "latitude"),  crs=4326)

yose_toads_sf
```

## Plot the toads with leaflet

We'll make the interactive map with leaflet, instead of tmap, because leaflet gives us more control over the popup windows.

The first step is to add a column to the sf object that contains the HTML code that will appear in the popup windows. We'll add this column using mutate, and the save the result as a new sf object:

```{r chunk09}
## Add a column containing HTML that will appear in the popup windows
yose_toads_popup_sf <- yose_toads_sf %>% 
  mutate(popup_html = paste0("<p><b>", common_name, "</b><br/>",
                             "<i>", scientific_name, "</i></p>",
                             "<p>Observed: ", datetime, "<br/>",
                             "User: ", user_login, "</p>",
                             "<p><img src='", image_url, "' style='width:100%;'/></p>")
  )
```

## Inspect the Popup HTML 

This is what the HTML code looks like for the first feature:

```{r chunk10}
yose_toads_popup_sf$popup_html[1]
```

Make the map

```{r chunk11}
leaflet(yose_toads_popup_sf) %>% 
  addTiles() %>% 
  addCircleMarkers(popup = ~popup_html, radius = 5)
```


## CHALLENGE: Download and map iNaturalist Observations for the taxon of your choice

Search for iNaturalist observations for your favorite Yosemite animal or plant by passing a Family, Genus or Species to `get_inat_obs()`. You could specify for example *Ursidae* (bears), *Sequoiadendron* (Sequoia), or *Lilium* (lilies). [Answer](https://bit.ly/3rETXmX)

```{r chunk12}
# Your answer here

```

## End

Congratulations, you've completed the Notebook! 

To view your Notebook at HTML, save it (again), then click the 'Preview' button in the RStudio toolbar.
