---
title: "R Syntax Review"
output: html_notebook
---

# About this R Notebook

R Notebooks are a 'flavor' of R markdown that combine plain text with R commands in code chunks. You can (and should!) edit them, and save it every few minutes like anything else!

Minimize the console. You won't need it because the output will appear below each code chunk.

To view a copy of your notebook as HTML, just hit Save, then Preview.

Keyboard shortcuts:

 - run the current line of R: *ctrl + enter*
 - run everything in the current code chunk: *ctrl + shift + enter*
 - insert a new code chunk: *ctrl + alt + i*


# Using R Like a Calculator

Compute the square root of 1000. [Answer](http://bit.ly/3vwTGE1)

```{r}

```

Write an expression that will return the area of a circle with radius = 2. [Answer](http://bit.ly/38OgpBY)

```{r}

```

Create a variable named x, and assign it a numeric value. Then write an expression that returns the diameter of a circle with radius = x. [Answer](http://bit.ly/3cDEHj7)

```{r}

```
Create two numeric variables and add them together. [Answer](http://bit.ly/3twiXML)

```{r}

```

What happens if you add a number variable to a character variable? [Answer](http://bit.ly/3eOgadJ)

```{r}

```

NOTE: YOU WON'T BE ABLE TO SAVE A R NOTEBOOK THAT HAS CODE CHUNKS THAT PRODUCE ERRORS.
COMMENT OUT ANY LINES OF CODE THAT GENERATE ERRORS BEFORE PROCEEDING.

# Syntax Rules

Fix this code (uncomment it first). [Answer](http://bit.ly/3rXcLx3)

```{r}
r <- 10

# circle_diameter <- R * 2

# circle_area >- pi * r ^ 2
```

# End

Congratulations on finishing the Notebook! Save it (again).

Click on the 'Preview' button in the RStudio toolbar to save a copy of your work as HTML.


