---
title: "R Syntax Review"
output: html_notebook
---

# Using R Like a Calculator

Compute the square root of 1000. [Answer](http://bit.ly/3vwTGE1)

```{r chunk01}
# Your answer here

```

Write an expression that will return the area of a circle with radius = 2. [Answer](http://bit.ly/38OgpBY)

Hint: circle area = pi * r squared

```{r chunk02}
# Your answer here

```

Create a variable named x, and assign it a numeric value. Then write an expression that returns the circumference of a circle with radius = x. [Answer](http://bit.ly/3cDEHj7)

Hint: circumference of a circle = 2 x pi x radius

```{r chunk03}
x <- 3

# circle circumference:

```

Create two numeric variables x and y, then add them together. [Answer](http://bit.ly/3twiXML)

```{r chunk04}
# Your answer here

```

What happens if you add a number variable to a character variable? [Answer](http://bit.ly/3eOgadJ)

```{r chunk05}
x <- 1
y <- "bandana"
# x + y
```

NOTE: YOU WON'T BE ABLE TO SAVE A R NOTEBOOK THAT HAS CODE CHUNKS THAT PRODUCE ERRORS.
COMMENT OUT ANY LINES OF CODE THAT GENERATE ERRORS BEFORE PROCEEDING.

# Syntax Rules

Fix this code (uncomment it first). [Answer](http://bit.ly/3rXcLx3)

```{r chunk06}
r <- 10

# circle_diameter <- R * 2

# circle_area >- pi * r ^ 2

```

# End

Congratulations on finishing the Notebook! Save it (again).

Click on the 'Preview' button in the RStudio toolbar to save a copy of your work as HTML.


