---
title: "Spatial Queries: Find Yosemite POIs in the Merced Watershed"
output: 
  html_notebook:
    toc: yes
    toc_float: yes
---

In this Notebook we'll use geoprocessing functions  from `sf` to identify Yosemite Points-of-Interest that fall within the **Merced watershed**. 

## Setup

Load the packages we'll need and set tmap mode to 'plot':

```{r chunk01}
library(sf)
library(tmap)
tmap_mode("plot")
```

Load `dplyr` and set name conflict preferences:

```{r chunk02}
library(dplyr)

## Load the conflicted package
library(conflicted)

# Set conflict preferences
conflict_prefer("filter", "dplyr", quiet = TRUE)
conflict_prefer("count", "dplyr", quiet = TRUE)
conflict_prefer("select", "dplyr", quiet = TRUE)
conflict_prefer("arrange", "dplyr", quiet = TRUE)
```

\

## Import the Watersheds

Start by importing the watersheds, grouping by the column HU (larger watershed units). Note how `group_by()` also aggregates (i.e., unions) spatial features associated with each unique value of the attribute:

```{r chunk03}
## Import the watersheds, grouping by HU
gpkg_watershd_fn <- "./data/yose_watersheds.gpkg"
yose_watersheds_hu <- st_read(gpkg_watershd_fn, layer="calw221") %>% 
  group_by(HU) %>% 
  summarise(HUNAME = first(HUNAME), NUM_WATERSHEDS = n())

## Plot results
tm_shape(yose_watersheds_hu) + 
  tm_polygons("MAP_COLORS", palette = "Pastel1") +
  tm_text("HUNAME", size = 0.7)
```

\

Next, pull out just the Merced River watershed and save it as a separate sf object:

```{r chunk04}
## Filter out just the Merced River watershid
merced_watershed <- yose_watersheds_hu %>% 
  filter(HUNAME == "MERCED RIVER")
```

\

## Import the Points-of-Interest

Import the POIs:

```{r chunk05}
## Import points of interest
yose_poi_utm <- st_read(dsn="./data", layer="yose_poi") %>% 
  select(OBJECTID, POINAME, POILABEL, POITYPE)
```

\

## Identify Points Which Intersect

Find out which POIs intersect the Merced River Watershed with `st_intersects()`:

```{r chunk06}
try(merced_poi <- yose_poi_utm %>% st_intersects(merced_watershed))
```

Oh no - **ERROR message**! Spatial querying requires features to be in the same CRS!

\

To fix this, we can project the HU watershed layer (which is in Albers) to match the POIs (which are UTM):

```{r chunk07}
merced_hu_utm <- merced_watershed %>% 
  st_transform(st_crs(yose_poi_utm))
```

\

Try the intersection test again:

```{r chunk08}
yose_poi_merced_mat <- yose_poi_utm %>% st_intersects(merced_hu_utm, sparse=FALSE)
head(yose_poi_merced_mat)
```

\

**CHALLENGE** How many points-of-interest fall within the Merced Rivershed watershed polygon? [Answer](https://bit.ly/3rEOnzn)

*Hint 1*: This is equivalent to asking how many TRUE values there are in the first column of `yose_poi_merced_mat`.

*Hint 2*: To get the first column of a matrix `x`, use `x[ , 1]`.

```{r chunk09}

```

\

## Subset the POIs that fall within the Merced River Watershed

To extract the POIs in the Merced watershed, we can feed the first column of `yose_poi_merced_mat` into  `filter()` (which expects TRUE/FALSE values):

```{r chunk10}
## Extract the points that intersect the watershed to a new sf object
merced_poi_utm <- yose_poi_utm %>% 
  filter(yose_poi_merced_mat[,1])
```

\

## Plot the Intersection

Plot to visually verify the results:

```{r chunk11}
## Plot
tm_shape(merced_hu_utm) +
  tm_polygons(col = "khaki") +
tm_shape(yose_poi_utm) +
  tm_dots(size = 0.2, col = "gray30") +
tm_shape(merced_poi_utm) +
  tm_dots(size = 0.2, col = "dodgerblue")
```

## End

Congratulations, you've completed another Notebook! 

To view your Notebook at HTML, save it (again), then click the 'Preview' button in the RStudio toolbar.

